---
title: "final"
author: "Genheylou Felisilda"
date: "2024-03-18"
output: pdf_document
---

# How to Create a 3D Population Density Map in R

#### **1. Install library Packages**:

You need to install the necessary packages. Run the following commands
in your R console. Try to install them one by one, it might library
restarting the R-session several times.

```{r}
install.packages("sf", dependencies=TRUE)
install.packages("tmap", dependencies=TRUE)
install.packages("mapview", dependencies=TRUE)
install.packages("stars", dependencies=TRUE)
install.packages("rayshader", dependencies=TRUE)
install.packages("MetBrewer", dependencies=TRUE)
install.packages("rayrender")
install.packages("extrafont", dependencies=TRUE)
install.packages("magick", dependencies=TRUE)
```

#### **2. Load Packages and Set Options**:

Load the library libraries and set the RGL options:

```{r}
options(rgl.useNULL = FALSE)
library(tidyverse)
library(sf)
library(tmap)
library(ggplot2)
library(mapview)
library(stars)
library(rayshader)
library(MetBrewer)
library(colorspace)
library(rayrender)
library(magick)
library(extrafont)
```

#### **3. Load and Transform Data**:

You'll need to load the population data and administrative boundaries
for the Philippines, transforming them into a suitable coordinate
system. The data is downloaded from Kontur Population.

```{r}
BB_hex <- st_read("data/kontur_population_BB_20231101.gpkg") %>% st_transform(3106)
BB_admin <- st_read("data/kontur_boundaries_BB_20230628.gpkg") %>% st_transform(3106)
```

#### **4. Check and Create Boundaries**:

Inspect the 'name_en' column and create the boundary for the
Philippines. Use the filter option to plot specific districts and
divisions on the map.

```{r}
distinct_names <- ph_admin %>% distinct(name_en)
print(distinct_names)
```

```{r}
# Creating BD Boundary
ph_boundary <- ph_admin %>%
  st_geometry %>%
  st_union %>%
  st_sf %>%
  st_make_valid()
```

#### **5. Plot Boundaries for Verification**:

Visualize the hex data and boundaries to ensure accuracy.

```{r}
names(ph_hex)
```

```{r}
 library(ggplot2)
ggplot(ph_hex) + 
  geom_sf(aes(fill = population), color = "gray66", linewidth = 0) + 
  geom_sf(data = ph_boundary, fill = NA, color = "black")
```

#### **6. Calculate Aspect Ratio**:

Determine the aspect ratio for the map based on the bounding box of the
boundary.

```{r}
# setting the ph boundary as a bounding box
bbox <- st_bbox(ph_boundary)

# finding the aspect ratio
bottom_left <- st_point(c(bbox[["xmin"]], bbox[["ymin"]])) %>%
  st_sfc(crs = 3106)
bottom_right <- st_point(c(bbox[["xmax"]], bbox[["ymin"]])) %>%
  st_sfc(crs = 3106)
top_left <- st_point(c(bbox[["xmin"]], bbox[["ymax"]])) %>%
  st_sfc(crs = 3106)
top_right <- st_point(c(bbox[["xmin"]], bbox[["ymax"]])) %>%
  st_sfc(crs = 3106)

width <- st_distance(bottom_left, bottom_right)
height <- st_distance(bottom_left, top_left)

if(width > height) {
  w_ratio = 1
  h_ratio = height / width
} else {
  h_ratio = 1.1
  w_ratio = width / height
}
```

#### 7. **Rasterize Population Data**:

Convert the population data into a raster format suitable for 3D
rendering.

-   For interactively checking the 3D plot setting the size low will
    help render in real time.

-   To improve the quality of the 3D image when saving, change the
    settings to a higher resolution.

```{r}
# convert to raster to convert to matrix
size = 3500

pop_raster <- st_rasterize(
  ph_hex,
  nx = floor(size * w_ratio) %>% as.numeric(),
  ny = floor(size * h_ratio) %>% as.numeric()
)

pop_matrix <- matrix(pop_raster$population,
                     nrow = floor(size * w_ratio),
                     ncol = floor(size * h_ratio))
```

#### **8. Define Color Palette**:

Select a color palette from the MetBrewer or RColorBrewer library and
customize it for your map.

```{r}
# Create color palette from MetBrewer Library
color <- MetBrewer::met.brewer(name="OKeeffe1", direction = -1)

tx <- grDevices::colorRampPalette(color, bias = 4.5)(256)
swatchplot(tx)
swatchplot(color)
```

#### **9. Render 3D Map**:

Use Rayshader to create a 3D representation of the population density.

```{r}
# Close any existing 3D plot before plotting another
rgl::close3d()

pop_matrix %>%
  height_shade(texture = tx) %>%
  plot_3d(heightmap = pop_matrix,
          #zscale = 250 / 4.5,
          zscale = 25,
          solid = F,
          shadowdepth = 0.8
          )

# Adjusting Camera Angle
render_camera(theta = 0,
              phi = 70,
              zoom = 0.55,
              fov = 100
)


# To interactively view the 3D plot
rgl::rglwidget()
```

#### **10. Render in high-quality and Save Image**:

Fine-tune the camera angle and render a high-quality image of the 3D
map.

```{r}
outfile <- glue::glue("Plots/Ph_final_2.png")

{
  start_time <- Sys.time()
  cat(crayon::cyan(start_time), "\n")
  if(!file.exists(outfile)) {
    png::writePNG(matrix(1), target = outfile)
  }
  
  render_highquality(
    filename = outfile,
    #interactive = F,
    lightdirection = 50, #Degree
    #lightaltitude = c(30, 80),
    #lightcolor = c(subset_colors[4], "white"),
    lightcolor = c("white", "white"),  # Set both lights to white
    lightintensity = c(600, 100),
    width = 1980,
    height = 1180,
    #width = 1000,
    #height = 1400,
    samples = 550
    #samples = 2
  )
  
  end_time <- Sys.time()
  diff <- end_time - start_time
  cat(crayon::cyan(diff), "\n")
}
```

\newpage

#### **11. Annotate the image**

You can add names and more details about your generated visualization.

```{r}
# ---------------------------Anotate
# Install and load the showtext package
install.packages("showtext")
library(showtext)
install.packages("extrafont")
library(extrafont)
font_import(pattern = "Philosopher")

pop_raster <- image_read("Plots/Ph_final.png")

text_color <- darken(color[1], .4)
swatchplot(text_color)


# Automatically enable font support
showtext_auto()

# Download and register the Philosopher font from Google Fonts
font_add_google("Philosopher", regular = "400", bold = "700")

pop_raster %>%
  image_annotate("Philippines",
                 gravity = "northeast",
                 location = "+50+50",
                 color = text_color,
                 size = 120,
                 font = "Philosopher",
                 weight = 700,
                 # degrees = 0,
  ) %>%
  image_annotate("POPULATION DENSITY MAP",
                 gravity = "northeast",
                 location = "+50+175",
                 color = text_color,
                 size = 28.5,
                 font = "Philosopher",  # Corrected font name
                 weight = 500,
                 # degrees = 0,
  ) %>%
  image_annotate("Visualization by: Insert Name \nData: Kontur Population 2023",
                 gravity = "southwest",
                 location = "+20+20",
                 color = alpha(text_color, .8),
                 font = "Philosopher",  # Corrected font name
                 size = 25,
                 # degrees = 0,
  ) %>%
  image_write("Plots/Annotated_plot_ph.png", format = "png", quality = 100)
```

