---
title: "final"
author: "Genheylou Felisilda"
date: "2024-03-18"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

# How to Create a 3D Population Density Map in R

#### **1. Install library Packages**:

You need to install the necessary packages. Run the following commands
in your R console. Try to install them one by one, it might library
restarting the R-session several times.

```{r}
install.packages("sf", dependencies=TRUE)
install.packages("tmap", dependencies=TRUE)
install.packages("mapview", dependencies=TRUE)
install.packages("stars", dependencies=TRUE)
install.packages("rayshader", dependencies=TRUE)
install.packages("MetBrewer", dependencies=TRUE)
install.packages("rayrender")
install.packages("extrafont", dependencies=TRUE)
install.packages("magick", dependencies=TRUE)
```

#### **2. Load Packages and Set Options**:

Load the library libraries and set the RGL options:

```{r}
options(rgl.useNULL = FALSE)
library(tidyverse)
library(sf)
library(tmap)
library(ggplot2)
library(mapview)
library(stars)
library(rayshader)
library(MetBrewer)
library(colorspace)
library(rayrender)
library(magick)
library(extrafont)
```

#### **3. Load and Transform Data**:

You'll need to load the population data and administrative boundaries
for the Philippines, transforming them into a suitable coordinate
system. The data is downloaded from Kontur Population.

```{r}
IT_hex <- st_read("C:/Program Files/RStudio/data/kontur_population_SI_20231101.gpkg") %>% st_transform(3106)
IT_admin <- st_read("C:/Program Files/RStudio/data/kontur_boundaries_SI_20230628.gpkg") %>% st_transform(3106)
```

#### **4. Check and Create Boundaries**: 

Inspect the 'name_en' column and create the boundary for the
Philippines. Use the filter option to plot specific districts and
divisions on the map.

```{r}
distinct_names <- IT_admin %>% distinct(name_en)
print(distinct_names)
```

```{r}
# Creating BD Boundary
IT_boundary <- IT_admin %>%
  st_geometry %>%
  st_union %>%
  st_sf %>%
  st_make_valid()
```

#### **5. Plot Boundaries for Verification**:

Visualize the hex data and boundaries to ensure accuracy.

```{r}
names(IT_hex)
```

```{r}
 library(ggplot2)
#ggplot(IT_hex) + 
#  geom_sf(aes(fill = population), color = "orange", linewidth = 0) + 
#  geom_sf(data = IT_boundary, fill = NA, color = "white")

# check the boundary plot
ggplot(IT_hex) +
  geom_sf(aes(fill = population),
          color = "skyblue",
          linewidth = 0) +
  geom_sf(
    data = IT_boundary,
    fill = NA,
    color = "black",
    linetype = "dashed",
    linewidth = 1
  )
```

#### **6. Calculate Aspect Ratio**:

Determine the aspect ratio for the map based on the bounding box of the
boundary.

```{r}
# setting the ph boundary as a bounding box
bbox <- st_bbox(IT_boundary)

# finding the aspect ratio
bottom_left <- st_point(c(bbox[["xmin"]], bbox[["ymin"]])) %>%
  st_sfc(crs = 3106)
bottom_right <- st_point(c(bbox[["xmax"]], bbox[["ymin"]])) %>%
  st_sfc(crs = 3106)
top_left <- st_point(c(bbox[["xmin"]], bbox[["ymax"]])) %>%
  st_sfc(crs = 3106)
top_right <- st_point(c(bbox[["xmin"]], bbox[["ymax"]])) %>%
  st_sfc(crs = 3106)

width <- st_distance(bottom_left, bottom_right)
height <- st_distance(bottom_left, top_left)

if(width > height) {
  w_ratio = 1
  h_ratio = height / width
} else {
  h_ratio = 1.1
  w_ratio = width / height
}
```

#### 7. **Rasterize Population Data**:

Convert the population data into a raster format suitable for 3D
rendering.

-   For interactively checking the 3D plot setting the size low will
    help render in real time.

-   To improve the quality of the 3D image when saving, change the
    settings to a higher resolution.

```{r}
# convert to raster to convert to matrix
size = 3500

pop_raster <- st_rasterize(
  IT_hex,
  nx = floor(size * w_ratio) %>% as.numeric(),
  ny = floor(size * h_ratio) %>% as.numeric()
)

pop_matrix <- matrix(pop_raster$population,
                     nrow = floor(size * w_ratio),
                     ncol = floor(size * h_ratio))
```

#### **8. Define Color Palette**:

Select a color palette from the MetBrewer or RColorBrewer library and
customize it for your map.

```{r}
# Create color palette from MetBrewer Library
color <- MetBrewer::met.brewer(name="Hiroshige", direction = -1)

tx <- grDevices::colorRampPalette(color, bias = 2.8)(256)
swatchplot(tx)
swatchplot(color)
```

#### **9. Render 3D Map**:

Use Rayshader to create a 3D representation of the population density.

```{r}
# Close any existing 3D plot before plotting another
rgl::close3d()

pop_matrix %>%
  height_shade(texture = tx) %>%
  plot_3d(heightmap = pop_matrix,
          zscale = 30,            
          solid = FALSE,
          shadowdepth = 0,
          theta = -20,
          phi = 20,
          zoom = .8,
          fov = 100
)
# To interactively view the 3D plot
rgl::rglwidget()
```

#### **10. Render in high-quality and Save Image**:

Fine-tune the camera angle and render a high-quality image of the 3D
map.

```{r}
outfile <- glue::glue("C:/Program Files/RStudio/data/SI_final_2[1].png")

{
  start_time <- Sys.time()
  cat(crayon::cyan(start_time), "\n")
  if(!file.exists(outfile)) {
    png::writePNG(matrix(1), target = outfile)
  }
  
  render_highquality(
    filename = outfile,
    interactive = FALSE,
    lightdirection = 55, #Degree
    lightaltitude = c(30, 80),
    #lightcolor = c(subset_colors[4], "white"),
    lightcolor = c("white", "white"),  # Set both lights to white
    lightintensity = c(600, 100),
    width = 1400,
    height = 1580,
    samples = 100
  )
  
  end_time <- Sys.time()
  diff <- end_time - start_time
  cat(crayon::cyan(diff), "\n")
}
```

\newpage

#### **11. Annotate the image**

You can add names and more details about your generated visualization.

```{r}
# Check if packages are installed, and install if necessary
if (!requireNamespace("showtext", quietly = TRUE)) {
  install.packages("showtext")
}
if (!requireNamespace("extrafont", quietly = TRUE)) {
  install.packages("extrafont")
}
if (!requireNamespace("magick", quietly = TRUE)) {
  install.packages("magick")
}

# Load required packages
library(showtext)
library(extrafont)
library(magick)

# Import fonts
font_import(pattern = "Philosopher")



# Automatically enable font support
showtext_auto()

# Load Google font
font_add_google("Philosopher", regular = "400", bold = "700")

# Read the SVG image
pop_raster <- image_read("C:/Program Files/RStudio/data/SI_final_2[1].png")

# Define text color

text_color <- "#1e466e" # Adjust as needed
text1_color <- "#376795"  

# Annotate the image
pop_raster %>%
  image_annotate("Slovenia",
                 gravity = "northeast",
                 location = "+50+50",
                 color = text_color,
                 size = 150,
                 font = "Philosopher",
                 weight = 800,
                 degrees = 0) %>%
  image_annotate("POPULATION DENSITY MAP",
                 gravity = "northeast",
                 location = "+50+230",
                 color = text_color,
                 size = 30,
                 font = "Philosopher",
                 weight = 500,
                 degrees = 0) %>%
  image_annotate("Visualization by: Culanggo | Felisilda | Casiño | Abainza \nData | Kontur Population 2023",
                 gravity = "southwest",
                 location = "+20+20",
                 color = alpha(text1_color, .8),
                 font = "Philosopher",
                 size = 22,
                 degrees = 0) %>%
  image_write("C:/Program Files/RStudio/data/eme_final_2[1].png", format = "png", quality = 100)
```
