---
title: "final"
author: "Genheylou Felisilda"
date: "2024-03-18"
output:
  word_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

## GROUP MEMBERS:

Felisilda \| Abainza \| Casiño \| Culanggo

#### **ERROR Found**

In this document, there is a recurring syntax error present in several
code blocks. The error arises from the absence of the language
identifier `{r}` within the code blocks, preventing the code from being
properly interpreted as R code.

In the original document, the language identifier `{r}` was written like
this:

``` r

# sample code
```

To resolve this issue, we made sure that all code blocks are preceded by
the language identifier `{r}` to indicate that they contain R code.

```{r}

# sample code

```

# How to Create a 3D Population Density Map in R

#### **1. Install library Packages**:

You need to install the necessary packages. Run the following commands
in your R console. Try to install them one by one, it might library
restarting the R-session several times.

```
install.packages("sf", dependencies=TRUE)
install.packages("tmap", dependencies=TRUE)
install.packages("mapview", dependencies=TRUE)
install.packages("stars", dependencies=TRUE)
install.packages("rayshader", dependencies=TRUE)
install.packages("MetBrewer", dependencies=TRUE)
install.packages("rayrender")
install.packages("extrafont", dependencies=TRUE)
install.packages("magick", dependencies=TRUE)
```

#### **2. Load Packages and Set Options**:

Load the library libraries and set the RGL options:

```{r}
options(rgl.useNULL = FALSE)
library(tidyverse)
library(sf)
library(tmap)
library(ggplot2)
library(mapview)
library(stars)
library(rayshader)
library(MetBrewer)
library(colorspace)
library(rayrender)
library(magick)
library(extrafont)
```

#### **3. Load and Transform Data**:

You'll need to load the population data and administrative boundaries
for the Philippines, transforming them into a suitable coordinate
system. The data is downloaded from Kontur Population.

```{r}

# population data of Myanmar
MM_hex <- st_read("data/kontur_population_MM_20231101.gpkg") %>% st_transform(3106)

# administrative boundaries of Myanmar
MM_admin <- st_read("data/kontur_boundaries_MM_20230628.gpkg") %>% st_transform(3106)

```

#### **4. Check and Create Boundaries**:

Inspect the 'name_en' column and create the boundary for the
Philippines. Use the filter option to plot specific districts and
divisions on the map.

```{r}
distinct_names <- MM_admin %>% distinct(name_en)
print(distinct_names)
```

```{r}
# Creating BD Boundary
MM_boundary <- MM_admin %>%
  st_geometry %>%
  st_union %>%
  st_sf %>%
  st_make_valid()
```

#### **5. Plot Boundaries for Verification**:

Visualize the hex data and boundaries to ensure accuracy.

```{r}
names(MM_hex)
```

## Revision No. 1

Upon reviewing the original code, we found that the representation of
the boundary lines in the plot lacked clarity and distinction. To
address this, we revised the code to enhance the styling of the boundary
lines in the plot.

We added the `linetype` and `linewidth` parameters in the `geom_sf`
function to include boundary lines in the plot.

```{r}
 library(ggplot2)

ggplot(MM_hex) +
  geom_sf(aes(fill = population),
          color = "orange",
          linewidth = 0) +
  geom_sf(
    data = MM_boundary,
    fill = NA,
    color = "black",
    
    # additional code for boundary lines
    
    linetype = "dashed",
    linewidth = .8
  )
```

#### **6. Calculate Aspect Ratio**:

Determine the aspect ratio for the map based on the bounding box of the
boundary.

```{r}
# setting the ph boundary as a bounding box
bbox <- st_bbox(MM_boundary)

# finding the aspect ratio
bottom_left <- st_point(c(bbox[["xmin"]], bbox[["ymin"]])) %>%
  st_sfc(crs = 3106)
bottom_right <- st_point(c(bbox[["xmax"]], bbox[["ymin"]])) %>%
  st_sfc(crs = 3106)
top_left <- st_point(c(bbox[["xmin"]], bbox[["ymax"]])) %>%
  st_sfc(crs = 3106)
top_right <- st_point(c(bbox[["xmin"]], bbox[["ymax"]])) %>%
  st_sfc(crs = 3106)

width <- st_distance(bottom_left, bottom_right)
height <- st_distance(bottom_left, top_left)

if(width > height) {
  w_ratio = 1
  h_ratio = height / width
} else {
  h_ratio = 1.1
  w_ratio = width / height
}
```

#### 7. **Rasterize Population Data**:

Convert the population data into a raster format suitable for 3D
rendering.

-   For interactively checking the 3D plot setting the size low will
    help render in real time.

-   To improve the quality of the 3D image when saving, change the
    settings to a higher resolution.

```{r}
# convert to raster to convert to matrix
size = 3500

pop_raster <- st_rasterize(
  MM_hex,
  nx = floor(size * w_ratio) %>% as.numeric(),
  ny = floor(size * h_ratio) %>% as.numeric()
)

pop_matrix <- matrix(pop_raster$population,
                     nrow = floor(size * w_ratio),
                     ncol = floor(size * h_ratio))
```

#### **8. Define Color Palette**:

Select a color palette from the MetBrewer or RColorBrewer library and
customize it for your map.

## Revision No. 2

n the original code, the color palette was generated with a
`bias of 4.5.` While the colors in the palette seemed fitting for our
visualization, we found that the transitions between colors were not as
distinct as we had hoped. To address this issue, we decided to change
the `bias from 4 to 4.5.`

This adjustment creates a color palette with more distinct boundaries
between colors and less smooth transitions compared to the original
palette. We expect that this change will lead to a perceptible
difference in the appearance of the generated color palette, resulting
in a better presentation of the population density of our chosen
country.

```{r}
# Create color palette from MetBrewer Library
color <- MetBrewer::met.brewer(name="Benedictus", direction = -1)

tx <- grDevices::colorRampPalette(color, bias = 4)(256) # new bias
swatchplot(tx)
swatchplot(color)

```

#### **9. Render 3D Map**:

Use Rayshader to create a 3D representation of the population density.

## Revision No. 3

In the original code, the `fov` parameter was included in the `plot_3d`
function. When we included it, we noticed that it affected the rendering
process of the 3D plot. Removing the FOV parameter allowed the rendering
process to proceed without the specific FOV setting. We made this
decision to optimize rendering efficiency, considering that specifying a
FOV may increase computational complexity or rendering time.

Although the `FOV` parameter was removed, the `plot_3d` function still
produces a reasonable output. The default angle leaves us satisfied, and
we believe that the inclusion of the FOV parameter does not
significantly affect the overall presentation of our input

```{r}
# Close any existing 3D plot before plotting another
rgl::close3d()

pop_matrix %>%
  height_shade(texture = tx) %>%
  plot_3d(heightmap = pop_matrix,
          zscale = 70,            
          solid = FALSE,
          shadowdepth = 0,
  )
render_camera(theta = -50, phi = 50, zoom = .7,)

# To interactively view the 3D plot
rgl::rglwidget()
```

#### **10. Render in high-quality and Save Image**:

Fine-tune the camera angle and render a high-quality image of the 3D
map.

## Revision No. 4

In this current chunk of code, multiple lines of code were originally
turned into comments, rendering them inexecutable. We resolved this
issue by removing the '\#' symbols to revert them back to executable
code.

However, after running this chunk multiple times, we encountered issues
generating high-quality files due to incompatible parameter values. The
original code produced unsuccessful outputs, such as all-black output,
all-black output with white dots, or a white dimension with a black
base. These issues stemmed from using
**`lightcolor = c("white", "white")`** instead of
**`lightcolor = c(pop_matrix[2], color)`**, which caused the 3D
population density map to not display any color.

To address this, we added **`sample_method = "sobol"`** to achieve a
more even coverage of space compared to purely random sequences.

We also made changes to the parameters in the **`render_highquality`**
function to resolve these issues. Refer to the comments in the code for
the specific changes made.

By incorporating these clarifications, the explanation becomes more
accessible and understandable for readers who may not be familiar with
the code or its context.

```
Sz# Define the output file path using the glue package
outfile <- glue::glue("C:/Users/HP PAVILION/Documents/data/MM_MAP[5].png")

{
  # Record the start time of the execution
  start_time <- Sys.time()
  
  # Print the start time in cyan color
  cat(crayon::cyan(start_time), "\n")
  
  if(!file.exists(outfile)) {
    png::writePNG(matrix(1), target = outfile)
  }
  
  render_highquality(
    filename = outfile,
    interactive = FALSE,     
    lightdirection = 100,    # originally 50,
    lightaltitude = c(70, 90),    # originally c(30,80)
    lightcolor = c(pop_matrix[2], color),   # originally c("white", "white")
    lightintensity = c(600, 100),
    samples = 450,   # originally 550
    sample_method = "sobol",
    width = 1080,    # originally 1980
    height = 920     # originally 1180
  )
  
  # Record the end time of the execution
  end_time <- Sys.time()
  
  # Calculate the difference between start and end time
  diff <- end_time - start_time
  
  # Print the execution time in cyan color
  cat(crayon::cyan(diff), "\n")
}

```

\newpage

#### **11. Annotate the image**

You can add names and more details about your generated visualization.

## Revision No. 5

We slightly enhanced the original code by revising these original lines:

`install.packages("showtext") library(showtext)`
`install.packages("extrafont") library(extrafont) font_import(pattern = "Philosopher")`

We transformed it into a code where it checks whether each package is
already installed before attempting to install it. This prevents
unnecessary installation processes, saving time and system resources.

While we changed some of the values for font customization, not much
were really revised. Some of the original parts of the code were still
used in this chunk. Any other changes made were not intently for a
different effect but just for preferences in customization.

```
# Check if packages are installed, and install if necessary
if (!requireNamespace("showtext", quietly = TRUE)) {
  install.packages("showtext")
}
if (!requireNamespace("extrafont", quietly = TRUE)) {
  install.packages("extrafont")
}
if (!requireNamespace("magick", quietly = TRUE)) {
  install.packages("magick")
}

# Load required packages
library(showtext)
library(extrafont)
library(magick)

# Import fonts
font_import(pattern = "Philosopher")



# Automatically enable font support
showtext_auto()

# Load Google font
font_add_google("Philosopher", regular = "400", bold = "700")

# Read the SVG image
pop_raster <- image_read("C:/Users/HP PAVILION/Documents/data/MM_MAP[3].png")

# Define text color

text_color <- "#1e466e" # Adjust as needed
text1_color <- "#376795"  

# Annotate the image
pop_raster %>%
  image_annotate("Myanmar",
                 gravity = "northeast",
                 location = "+50+50",
                 color = text_color,
                 size = 150,
                 font = "Philosopher",
                 weight = 800,
                 degrees = 0) %>%
  
  image_annotate("POPULATION DENSITY MAP",
                 gravity = "northeast",
                 location = "+50+230",
                 color = text_color,
                 size = 30,
                 font = "Philosopher",
                 weight = 500,
                 degrees = 0) %>%
  
  image_annotate("Visualization by: Culanggo | Felisilda | Casiño | Abainza \nData @ Kontur Population 2023",
                 gravity = "southwest",
                 location = "+20+20",
                 color = alpha(text1_color, .8),
                 font = "Philosopher",
                 size = 22,
                 degrees = 0) %>%
  
  image_write("C:/Users/HP PAVILION/Documents/data/MM_MAP[3](ANNOTATED).png", format = "png", quality = 100)
```
